Tidy data structure to support exploration and modeling of temporal data

### 2019-01-04 main.Rmd
a bit more on implicit na

The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modelling. Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks. With the pipe operator `%>%`, read from left to right, and omit intermediate values.

@@

XXX Put comments in above each paragraph, with the topic of each paragraph. its not clear how this section flows yet. the topics would help.

The time-based pipeline shepards raw temporal data through to time series analysis, and plots. It is possible and recommended to check for identical entries of key and index before analysis. Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the exploration stage.

Time series models typically assume that the input is a complete and regularly-spaced series, which most temporal data can barely satisfy. Temporal data becomes more granular in time resolution and more disaggregated to individual level. This inevitably goes with implicit missing values and noisiness. Data needs transformation to some extent for modelling. A suite of verbs are introduced to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, as well as transformed for visualization. The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood. Attention has been paid to error handling. If a tsibble cannot be maintained in the output of a pipeline module, for example the index is dropped by aggregation, an error informs users of the problem and suggests alternatives. This avoids negatively surprising users and reminds them of time context.

@@

@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony. This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems. Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, known as "tidyverts".

>>>>

The tidy data abstraction lays a pipeline infrastructure for data analysis modules of transformation, visualization and modelling. Each module communicates between each other, requiring tidy input, producing tidy output, chaining a series of operations together to accomplish the analytic tasks.

@@

The time-based pipeline shepards raw temporal data through to time series analysis, and plots. It is possible and recommended to check for identical entries of key and index before analysis. Duplicates signal the data quality issue, which would likely affect succeeding analyses and hence decision making. Analysts are encouraged to gaze at data early and reason about the process of data cleaning. When the data meets the tsibble standard, it flows neatly into the analysis stage.

Many time operations like lag/lead and time series models, assume an intact vector input ordered in time. To prevent inviting these errors into the analysis, it is a good practice to check and inspect any time gaps of the tsibble in the first place. Several tools are provided to understand and tackle implicit missing values in: (1) `has_gaps()` checks if there exists time gaps; (2) `scan_gaps()` reports all implicit missing observations; (3) `count_gaps()` summarises the time ranges that are absent from the data; (4) `fill_gaps()` turns them into explicit ones, along with imputing by values or functions.

Data needs transformation to some extent for modelling. A suite of verbs are introduced to flatten the lumpy path from temporal data to an object that can be directly modeled in the tsibble framework, as well as transformed for visualization. The principle that underpins most verbs is a tsibble in and a tsibble out, thereby striving to retain a valid tsibble by automating index and key updates under the hood. Attention has been paid to error handling. If a tsibble cannot be maintained in the output of a pipeline module, for example the index is dropped by aggregation, an error informs users of the problem and suggests alternatives. This avoids negatively surprising users and reminds them of time context.

@@

@eopl asserted "No matter how complex and polished the individual operations are, it is often the quality of the glue that most directly determines the power of the system." Each verb works with other transformation family members in harmony. This set of verbs can result in many combinations to prepare tsibble for a broad range of visualization and modeling problems. Chaining operations is achieved with the pipe operator `%>%` introduced in the **magrittr** package. A sequence of functions can be composed in a way that can be naturally read from left to right, which incredibly increase the readability of the code. It consequently generates a block of code without worrying naming intermediate values. Most importantly, the ecosystem for tidy time series analysis has been undertaking on the basis of tsibble in R, known as "tidyverts".
